---
title: "lab16"
author: "Ziyuan_Han"
date: "11/19/2021"
output:
  pdf_document: default
  html_document: default
---


#Differential expression gene analysis
```{r}
library(DESeq2)
metaFile <- "GSE37704_metadata.csv"
countFile <- "GSE37704_featurecounts.csv"
# Import metadata and take a peak
colData = read.csv(metaFile, row.names=1)
head(colData)
```
## Import countdata
```{r}
countData = read.csv(countFile, row.names=1)
head(countData)
```
# Note we need to remove the odd first $length col
```{r}
countData <- as.matrix(countData[,-1])
head(countData)
```
# Filter count data where you have 0 read count across all samples.
Q. Complete the code below to filter countData to exclude genes (i.e. rows) where we have 0 read count across all samples (i.e. columns).
Tip: What will rowSums() of countData return and how could you use it in this context?
```{r}
countData = countData[-which(rowSums(countData)==0),]
head(countData)
```
#PCA
```{r}
pca <- prcomp(t(countData))
summary(pca)
plot(pca$x[,1:2])
library(ggplot2)
x<-as.data.frame(pca$x)
x$condition <- colData$condition
ggplot(x) +
  aes(PC1,PC2,col=condition)+
  geom_point()
```
#Running DESeq2
```{r}
dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)
dds
res = results(dds)
summary(res)
```

#Volcano plot
```{r}
plot( res$log2FoldChange, -log(res$padj) )
# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
inds <- (abs(res$pvalue) <0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```
#Adding gene annotation
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
columns(org.Hs.eg.db)
res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(res), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="GENENAME",
                    multiVals="first")

head(res, 10)
```
#Q. Finally for this section let's reorder these results by adjusted p-value and save them to a CSV file in your current project directory.
```{r}
res = res[order(res$pvalue),]
write.csv(res,file="deseq_results.csv")
```

#Pathway analysis

#BiocManager::install( c("pathview", "gage", "gageData") )

# For old vesrsions of R only (R < 3.5.0)!
#source("http://bioconductor.org/biocLite.R")
#biocLite( c("pathview", "gage", "gageData") )
```{r}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
attributes(keggres)
# Look at the first few down (less) pathways
head(keggres$less)
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```
# A different PDF based output of the same data
```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```
## Focus on top 5 upregulated pathways here for demo purposes only
```{r}
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```

#Gene Ontology
```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```
#Reactome Analysis
```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
write.table(sig_genes, file="significant_genes.txt", row.names=FALSE, col.names=FALSE, quote=FALSE)
```
