---
title: "lab10"
author: "Ziyuan_Han"
date: "10/31/2021"
output:
  word_document: default
  pdf_document: default
---

#Class 10: Halloween Mini-Project

##Exploratory Analysis of Halloween Candy

#1. Importing candy data
#https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-powerranking/candy-data.csv
```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy = read.csv(candy_file, row.names=1) 
head(candy)
```
#Q1. How many different candy types are in this dataset?

85
```{r}

nrow(candy)
```
#Q2. How many fruity candy types are in the dataset? The functions dim(), nrow(), table() and sum() may be useful for answering the first 2 questions.

38
```{r}
nrow(candy[candy$fruity == 1,])
```

#What is your favorate candy?
```{r}
candy["Twix", ]$winpercent
```
# What is your favorite candy in the dataset and what is it’s winpercent value?
ReeseÕs Peanut Butter cup: winpercent is 84.18029
```{r}
candy["ReeseÕs Peanut Butter cup",]$winpercent
```
#What is the winpercent value for “Kit Kat”?
```{r}
candy["Kit Kat",]$winpercent
```

#What is the winpercent value for “Tootsie Roll Snack Bars
```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
#install.packages("devtools") 
#devtools::install_github("ropensci/skimr")
library("skimr")
skim(candy)

```
#Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
column 12 is inscale. So we have to scale the data when doing PCA otherwise this parameter is going to dominant over the rest.

#Q7. What do you think a zero and one represent for the candy$chocolate column?
0 and 1 represent boolean values False and True. Indicating the candy contains cholocate or not.

#Q8. Plot a histogram of winpercent values
```{r}
library(ggplot2) 
data = candy 
data$type = rownames(data) 
ggplot(data, aes(x=winpercent)) + geom_histogram()
```
#Q9. Is the distribution of winpercent values symmetrical?

Yes 

#Q10. Is the center of the distribution above or below 50%?

obove 50% 

#Q11. On average is chocolate candy higher or lower ranked than fruit candy? 
cholocate candy rank higher than fruit candy

```{r}
print(mean(candy$winpercent[as.logical(candy$chocolate)]))
choc=data[data$chocolate == 1,] 
ggplot(choc, aes(x=winpercent)) + 
  geom_histogram() + xlim(0,100)
fruit=data[data$fruity == 1,] 
ggplot(fruit, aes(x=winpercent)) + 
  geom_histogram() + xlim(0,100)
```
#Q12. Is this difference statistically significant? 
p-val of T-test is less than 0.05, indicating there is statistical significance between preferences for chocolate and fruity candy.

```{r}
choc = candy$winpercent[as.logical(candy$chocolate)] 
fruit = candy$winpercent[as.logical(candy$fruity)] 
t.test(choc, fruit)
```
#Overall Candy Rankings
```{r}
library(dplyr)
candy %>% arrange(winpercent) %>% head(5)
head(candy[order(candy$winpercent),], n=5)
```
#Q13. What are the five least liked candy types in this set?
```{r}
candy %>% arrange(winpercent) %>% head(5)
```
#Q14. What are the top 5 all time favorite candy types out of this set?
```{r}
candy %>% arrange(desc(winpercent)) %>% head(5)
```
#Q15. Make a first barplot of candy ranking based on winpercent values. HINT: Use the aes(winpercent, rownames(candy)) for your first ggplot like so:
```{r}
library(ggplot2) 
ggplot(candy) +
aes(x=winpercent, y=rownames(candy)) + 
  geom_col()
```
#Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent? HINT: You can use aes(winpercent, reorder(rownames(candy),winpercent)) to improve your plot.
```{r}
ggplot(candy) +
aes(winpercent, reorder(rownames(candy),winpercent)) + 
  geom_col()

ggplot(candy) +
aes(winpercent, reorder(rownames(candy),winpercent)) + 
  geom_col()
```
#Q17. What is the worst ranked chocolate candy?
sixlets
```{r}
ggplot(candy) +
aes(winpercent, reorder(rownames(candy),winpercent)) 
```

#Q18. What is the best ranked fruity candy?
starburst

#Taking a look at pricepercent
```{r}
library(ggrepel)
aes(winpercent, pricepercent, label=rownames(candy)) + 
  geom_text_repel(size=3.3, max.overlaps = 5)
```
#Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
HersheyÕs Krackel

#Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
Nik L Nip
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE) 
head( candy[ord,c(11,12)], n=5 )
```
#5 Exploring the correlation structure Now that we’ve explored the dataset a little, we’ll see how the variables interact with one another. We’ll use correlation and view the results with the corrplot package to plot a correlation matrix.
```{r}
library(corrplot)
cij <- cor(candy) 
corrplot(cij)
```
#Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
fruity and chocolate 
#Q23. Similarly, what two variables are most positively correlated?
winpercent and chocolate 

#Principal Component Analysis
```{r}
pca <- prcomp(candy, scale.=TRUE) 
summary(pca)
plot(pca$x[,1:2])
plot(pca$x[,1:2], pch=16)
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) +
aes(x=PC1, y=PC2, size=winpercent/100, text=rownames(my_data), label=rownames(my_data)) 
p
library(ggrepel)
p + geom_text_repel(size=3.3, max.overlaps = 7)
theme(legend.position = "none") + labs(title="Halloween Candy PCA Space",subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red)",caption="Data from 538")
par(mar=c(8,4,2,2)) 
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
#Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you? HINT. pluribus means the candy comes in a bag or box of multiple candies.

fruity, hard, pluribus. 
yes
