---
title: "lab17"
author: "Ziyuan_Han"
date: "12/1/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

# Import vaccination data
```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```
##Q1. What column details the total number of people fully vaccinated? 
age12_plus_population and age5_plus_population
##Q2. What column details the Zip code tabulation area? 
column 2 "zip_code_tabulation_area"
##Q3. What is the earliest date in this dataset? 
2021.1.5
##Q4. What is the latest date in this dataset?
2021.11.30
```{r}
head(sort(vax$as_of_date,decreasing=TRUE))
```
```{r}
#skimr::skim(vax)
```
##Q5. How many numeric columns are in this dataset? 
9
##Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 
8472
##Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?
10%
```{r}
8472/84672 
```
#Working with dates
```{r}
library(lubridate)
today
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
#today() - vax$as_of_date[1]
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
length(unique(vax$as_of_date))
```
##How many days have passed since the last update of the dataset? 
329 days
## How many unique dates are in the dataset (i.e. how many different dates are detailed)? 
48
#Working with zip codes
```{r}
#install.packages("zipcodeR")
library(zipcodeR)
geocode_zip('92037')
zip_distance('92037','92109')
reverse_zipcode(c('92037', "92109") )
zipdata <- reverse_zipcode(vax$zip_code_tabulation_area )
# Subset to San Diego county only areas
#install.packages("dplyr")
library(dplyr)
#Focus on San Deigo
sd <- filter(vax, county == "San Diego")

nrow(sd)

sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```
#Q11. How many distinct zip codes are listed for San Diego County? 
76
```{r}
length(unique(sd.10$zip_code_tabulation_area))
```

#Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 
92154  
#Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”? 
0.6802456
```{r}
sd.20211109 <- filter(sd.10,as_of_date=="2021-11-09")
mean(sd.20211109$percent_of_population_fully_vaccinated)
```


#Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?
```{r}
hist(sd.20211109$percent_of_population_fully_vaccinated)
```


#focus on UCSD/La Jolla
```{r}
ucsd <- filter(sd,zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

#Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:
```{r}
library(ggplot2)
p<-ggplot() +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",main="Vaccination rate for La Jolla CA92109")
p
```
# Subset to all CA areas with a population as large as 92037
```{r}
library(dplyr)
vax.36 <- filter(vax, age5_plus_population > 36144 &as_of_date == "2021-11-16")
```
#Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?
```{r}
mean<-mean(vax.36$percent_of_population_fully_vaccinated)
p<- p + 
  geom_hline(yintercept=mean)
p
```
sessionInfo()

