---
title: "lab09PCA"
author: "Ziyuan_Han"
date: "10/27/2021"
output: pdf_document
---
#Get the dataset into the project
```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, row.names=1)
head(wisc.df)
```
#visualize the data 
```{r}
wisc.data <- wisc.df[,-1]
# Create diagnosis vector for later 
diagnosis <- as.factor(wisc.df$diagnosis)
```
#Q1. How many observations are in this dataset?
569 observations
```{r}
nrow(wisc.df)
```
Q2. How many of the observations have a malignant diagnosis?
212 observations have the malignant diagnosis
```{r}
length(which(diagnosis=="M"))
```

Q3. How many variables/features in the data are suffixed with _mean?
10 
```{r}
colname<-colnames(wisc.df)
length(grep("_mean",colname))
```

#PCA analysis 
##Check column means and standard deviations
```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)
```
# Perform PCA on wisc.data by completing the following code
```{r}
wisc.pr <- prcomp(wisc.data,scale.=TRUE)
summary(wisc.pr)
plot(wisc.pr)
```
#Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
44.27% of the variance is captured 

#Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
3 PCs are required to describe 72.6% of the original variance

#Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
7 PCs are required to describe at least 90% of the original variance in the data 

# Interpreting PCA results 
#Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?
It has too many labels and looks like messy. It's difficult to understand because the annotation is not clear.
```{r}
library(ggplot2)
biplot(wisc.pr)
```
# Scatter plot observations by components 1 and 2
```{r}
plot(wisc.pr$x, col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```
#Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
the PC1 and PC2 plot look similar to PC1 and PC3 plot, which indicate the first two PCs can explain most variance in the data
```{r}
plot(wisc.pr$x[,1:3], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```
# Create a data.frame for ggplot
```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis

# Load the ggplot2 package
library(ggplot2)

# Make a scatter plot colored by diagnosis
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) +
  geom_point()
```
#Variance Explained
# Calculate variance of each component
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```
# Alternative scree plot of the same data, note data driven y-axis
```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```
## ggplot based graph
```{r}
#install.packages("factoextra")
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```

#Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?
-0.26085376 
```{r}
wisc.pr$rotation[,1]
```
#Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?
5 PCs
```{r}
summary(wisc.pr)
```

#Hierarchical clustering
# Scale the wisc.data data using the "scale()" function
```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist,method="complete")
```
#Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?
around 20
```{r}
plot(wisc.hclust)
abline(h=20, col="red", lty=2)
```
#Selecting number of clusters
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust,k=4)
table(wisc.hclust.clusters, diagnosis)
```
#Q12. Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10?
5 is a better cluster match 
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust,k=2)
table(wisc.hclust.clusters, diagnosis)
wisc.hclust.clusters <- cutree(wisc.hclust,k=10)
table(wisc.hclust.clusters, diagnosis)
wisc.hclust.clusters <- cutree(wisc.hclust,k=5)
table(wisc.hclust.clusters, diagnosis)
```
#Using different methods
#Q13. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.
I love ward.D2 best because it gave me fewer clusters but more clear grouping 
```{r}
wisc.hclust.single <- hclust(data.dist,method="single")
plot(wisc.hclust.single)
wisc.hclust.complete <- hclust(data.dist,method="complete")
plot(wisc.hclust.complete)
wisc.hclust.average <- hclust(data.dist,method="average")
plot(wisc.hclust.average)
wisc.hclust.ward <- hclust(data.dist,method="ward.D2")
plot(wisc.hclust.ward)
```
#4. OPTIONAL: K-means clustering
#Q14. How well does k-means separate the two diagnoses? How does it compare to your hclust results?
Personally, I think k-means separate the two data better since the different clusters are more separated.

```{r}
wisc.km <- kmeans(scale(wisc.data), centers= 2, nstart= 20)
table(wisc.km$cluster, diagnosis)
table(wisc.km$cluster, wisc.hclust.clusters)
```
#Combining methods
```{r}
wisc.pr.hclust <-hclust(dist(wisc.pr$x[,1:7]),method="ward.D2")
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, diagnosis)
plot(wisc.pr$x[,1:2], col=grps)
plot(wisc.pr$x[,1:2], col=diagnosis)
g <- as.factor(grps)
levels(g)
g <- relevel(g,2)
levels(g)
# Plot using our re-ordered factor 
plot(wisc.pr$x[,1:2], col=g)
```
#Q15. How well does the newly created model with four clusters separate out the two diagnoses?
The newly created model can separate the clusters well
```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
table(wisc.pr.hclust.clusters, diagnosis)
```
#Q16. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses
Hierarchical clustering models perform better since it separates clusters into more groups and more separate.
```{r}
table(wisc.km$cluster, diagnosis)
table(wisc.hclust.clusters, diagnosis)
```
#Q17. Which of your analysis procedures resulted in a clustering model with the best specificity? How about sensitivity?
Variance test aims to increase sensitivity while the clustering process improves specificity.

#Prediction
#url <- "new_samples.csv"
#Q18. Which of these new patients should we prioritize for follow up based on your results?
Patient 2 needs more follow up
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```
sessionInfo()


