---
title: "class06"
author: "Ziyuan_Han"
date: "10/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## quick R Markdown intro
We can write text of course just like any file, we can **style the text to be bold** or *italic*

Do:
-this
-that
-and another thing

this is new text (two whitespaces to create a new line)  
and this is a new line  
---
we can include some code:
```{r}
# this is a comment
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```
short cut for adding code chunck
option + command + I (for R code)

**Q1.** Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]
```{r}
grade <- function(score){
  if (sum( is.na( data ) ) > 0){
    
  }
  del_min = score[-match(min(score), score)]
  mean(del_min)
}

avg_del_min = grade(student3)
avg_del_min

```
First, find the lowest score using **min()**, and the **which.min()** function to find where it is (ie. the position in the vector)
```{r}
which.min(student1)
```

```{r}
student1[-which.min(student1)]
```
Now I can call the **mean()** function to get the average.
```{r}
mean(student1[-which.min(student1)])
```

```{r}
which.min(student2)
mean(student2,na.rm=TRUE)
```






